---
title: "HW7-WQ"
author: "Wei Qiu"
date: "2022-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use the MySQL database imdb hosted at scidb.smith.edu (usersname: sds192,
password: DSismfc@S) to answer the following questions.

1. Write a SQL query to create a list of all movies that Kevin Bacon has ever been in and convert this to a data frame object in R/Python.

```{r}
library(tidyverse)
library(RMySQL)
db <- dbConnect(
  MySQL(),
  host = "scidb.smith.edu",
  user = "sds192",
  password = "DSismfc@S",
  dbname = "imdb" # one database for movie and TV
)
```


```{r}
# list of tables
dbListTables(db)
```


```{r}
# I will need person_info table, name table, cast_info table, and title.The connection between person_info and cast_info is person_id, and the relationship between cast_info and title is movie_id. person_info and name table have person_id. 

# Strategy: combine name with person_info by person_id, and then combine them with cast_info by person_id, and then combine them with title with movie_id.  

# relationship among tables: https://www.researchgate.net/figure/Complete-schema-of-the-IMDb-database-with-8-main-relations-movie-person-genre_fig1_348079657

# Double check the colnames

#person_info <- tbl(db,"person_info")
#person_info %>% colnames()
#person_info %>% head(5)

#name <- tbl(db,'name')
#name %>% colnames()
#name %>% head(5)

#cast_info <- tbl(db,'cast_info')
#cast_info %>% colnames()
#cast_info %>% head(5)


#title <- tbl(db,'title')
#title %>% colnames()
#title <- title %>% head(20)
```


```{r}

# Strategy: combine name (n.id) with cast_info by person_id, and then combine them with title with movie_id.


```

```{r}
#Method 1

q <- "SELECT DISTINCT t.title, ci.movie_id,t.production_year
FROM name n
JOIN cast_info ci ON ci.person_id = n.id
JOIN title t ON t.id = ci.movie_id
WHERE n.name='Bacon, Kevin'
AND t.kind_id=1
ORDER BY production_year
DESC;"

result = dbSendQuery(db, q) 
fetch(result)
```


```{r}
# Method 2.
q <- "SELECT DISTINCT t.title, ci.movie_id,t.production_year
FROM name n, cast_info ci,title t
WHERE ci.person_id = n.id
AND t.id = ci.movie_id
AND n.name='Bacon, Kevin'
AND t.kind_id=1
ORDER BY production_year
DESC;"

result = dbSendQuery(db, q) 
fetch(result)
```


2. Find a list of every actor/actress who has appeared in a movie with Kevin Bacon.

```{r}
q <- "SELECT DISTINCT ci.person_id, n.name
FROM name n
JOIN cast_info ci ON ci.person_id = n.id
JOIN title t ON t.id = ci.movie_id
WHERE t.kind_id=1
AND (ci.role_id=1 OR ci.role_id=2)
AND ci.movie_id IN 
(SELECT DISTINCT ci.movie_id
FROM name n
JOIN cast_info ci ON ci.person_id = n.id
JOIN title t ON t.id = ci.movie_id
WHERE n.name='Bacon, Kevin'
AND t.kind_id=1
);"
result = dbSendQuery(db, q) 
fetch(result,-1)
```

# disconnect to release memory
DBI::dbDisconnect(db)

# Python homework

https://github.com/jamesqiu2021/DSCI-401-QW/blob/main/Copy_of_hw7_qw.ipynb

