---
title: "Exam1 QW"
author: "Wei Qiu"
date: "2022-10-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load the cereal data and use it to answer the following questions. (Note: For loops are not an acceptable solution for any of these questions.)

```{r}
library(tidyverse)
library (corrr)
cereal <- read.csv("https://raw.githubusercontent.com/gjm112/LoyolaTeaching/main/cereal.csv")
manufactures <- read.csv("https://raw.githubusercontent.com/gjm112/LoyolaTeaching/main/cereal_mfr.csv")

```

1(a) (5 points) Merge the manufacturer full name onto the cereal data set.

```{r}
cereal_merge <- cereal %>% inner_join(manufactures,by=c("mfr"="mfr"))
cereal_merge %>% head(5)
```

1(b) (5 points) For this problem, we will only consider low sodium cereals" defined as cereals with less than or equal to 200mg of sodium.
Create a table for the low sodium cereals that shows the mean calories, mean protein, and mean fat content by the full name of the
manufacturer.

```{r}
cereal_merge1 <- cereal_merge %>% select(mfr_full,calories, protein, fat,sodium) %>% filter(sodium <= 200) %>% group_by(mfr_full) %>% summarize(mean_sodium = mean(sodium),mean_calories = mean(calories),mean_protein = mean(protein),mean_fat = mean(fat))
cereal_merge1
```
(c) (5 points) Which 5 cereals have the highest calories per cup (i.e calories / cups)?

```{r}
cereal_merge5 <- cereal_merge %>% mutate(calories_per_cup = calories/cups) %>% select(name,calories_per_cup) %>%  arrange(desc(calories_per_cup)) 
cereal_merge5 %>% head(5)
```
(d) (5 points) For each manufacturer, which of their cereals has the lowest calories per cup (i.e. calories divided by cups)?

```{r}
cereal_merge2 <- cereal_merge %>% mutate(calories_per_cup = calories/cups) %>%  group_by(mfr_full) %>% summarise(min_calories_cup=min(calories_per_cup))%>% inner_join(cereal_merge5,by=c("min_calories_cup"="calories_per_cup"))
cereal_merge2 %>% head(5)
```
(e) (10 points) Create a scatterplot of sugars vs rating for cereals manufactured by General Mills, Kellogg's, and Post with color used to
represent each of the different manufacturers. Add a trend line using
loess through each of the groups separately. Comment on the rela-
relationship between sugars and rating for these manufacturers (target
figure shown below).

```{r}
#generate the new dataframe
cereal_merge3 <- cereal_merge %>% select(mfr_full,sugars,rating) %>% filter(mfr_full=="General Mills"|mfr_full=="Post"|mfr_full=="Kelloggs") 
cereal_merge3 %>% head(10)

#plots
plots <- ggplot(data=cereal_merge3,aes(x=sugars,y=rating))+geom_point(aes(color = mfr_full))+geom_smooth(aes(color = mfr_full),method = "loess", se = TRUE)
plots

# comments: The level of sugars in the cereals was negatively correlated with the rating from all these manufactures. 
```

(f) (10 points) Find the correlation between protein and potassium for
each of the shelves (the correlation function in R is "cor").


```{r}
cereal_merge %>% select(shelf,protein,potass) %>% group_by(shelf) %>% group_modify(~correlate(.x),keep=FALSE)
```

(g) (10 points) Recreate the table presented below. The rows are the
shelf 1, 2, and 3 and each column is a quantile of the distribution of
calories for each shelf. So for example, the 80-th percentile of calories
for cereal found on shelf 3 is 130.
## # A tibble: 3 x 12
## # Groups: shelf [3]
## shelf 0 0.1 0.2 0.3 0.4 0.5 0.6 0.7 0.8 0.9 1
## <int> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
## 1 1 80 90 98 100 100 105 110 110 110 110 110
## 2 2 90 100 100 110 110 110 110 110 120 120 120
## 3 3 50 70 90 100 100 110 110 120 130 140 160

```{r}
#generate narrow data
g_narrow <- cereal_merge %>% group_by(shelf) %>%  summarise(quantile = scales::percent(c(0,.1,.2,.3,.4,.5,.6,.7,.8,.9,1.0)),
            calories = quantile(calories, c(0,.1,.2,.3,.4,.5,.6,.7,.8,.9,1.0)))
g_narrow

#pivot wider

g_wider <- g_narrow %>% pivot_wider(names_from = quantile,values_from = calories) 
g_wider
```


2. Use the data found here about food inspection found here to answer the following questions. (Note: Remove observations with license number 0 be-
fore beginning!) https://www.dropbox.com/s/ra8s2tonbl037m2/inspections clean.csv?dl=0

```{r}
inspection <- read.csv("inspections_clean.csv")
inspection %>% head(1)
```

(a) (10 points) What is the pass rate by inspection type?


```{r}
# I did not include pass w/conditions as suggested in 2c. 
inspection_test <- inspection
inspection_test %>% mutate(pass = if_else(results=="Pass",1,0)) %>% select (inspection_type,pass) %>%  group_by(inspection_type) %>% summarise(pass_rate = sum(pass) / nrow(inspection_test))
```
(b) (10 points) Find all the businesses that have ever had a failed inspection. Using only these businesses, create a contingency table of all the inspections for these businesses of risk type by inspection type.
It should look like the table below:
## # A tibble: 4 x 4
## # Groups: risk [4]
## risk canvass complaint license
## <chr> <int> <int> <int>
## 1 Risk 1 (High) 60151 14355 10646
## 2 Risk 2 (Medium) 10101 4556 2942
## 3 Risk 3 (Low) 1901 875 1854
## 4 <NA> NA NA 1


```{r}
#Find all the businesses that have ever had a failed inspection
inspection_test4 <-inspection %>% filter(results=="Fail") %>% select(dba_name) %>% distinct()
inspection_test4 %>% head(5)
```

```{r}
inspection_test5 <- inspection  %>% inner_join(inspection_test4,by=c("dba_name"="dba_name")) %>% group_by(risk,inspection_type) %>% summarise(n=length(risk)) %>% pivot_wider(names_from = inspection_type,values_from = n)
inspection_test5
```

(c) (10 points) What is the restaurant with the largest number of inspections that has passed (do NOT include \passed with conditions")
ALL of their inspections and how many inspections did they have?
That is which restaurant had the most inspections without any failure?


```{r}
#find all business failed or passed with conditions
inspection_fail <- inspection %>% filter(facility_type=="restaurant") %>% filter(results!="Pass") %>% select(dba_name) %>% distinct() 
inspection_fail %>% head(5)
```

```{r}
inspection %>% filter(facility_type=="restaurant") %>% anti_join(inspection_fail,by="dba_name") %>% filter(results=="Pass") %>% group_by(dba_name) %>% summarize(inspection_number=length(results)) %>% arrange(desc(inspection_number)) %>% head(5)
```
#MANNY'S CAFE MIDWAY had the most inspections without any failure.


```{r}
#double check
inspection %>% select(dba_name,results) %>% filter(dba_name=="MANNY'S CAFE MIDWAY") %>% head(5)
```

(d) (10 points) Make a stacked bar chart showing the number of tests for
grocery stores, restaurants, and schools with each bar stratified by
results (target plot is displayed below).

```{r}
inspection %>% filter(facility_type %in% c("restaurant","school","grocery store")) %>% group_by(results,facility_type) %>% summarise(number_of_tests=length(inspection_id)) %>% ggplot(aes(x = facility_type, y = number_of_tests, fill = results)) + 
  geom_bar(stat = "identity")
```


(e) (10 points) Make side by side boxplots of the distribution of the
number of tests establishments have performed on them stratified by
whether or not an establishment has ever failed a failed test (target
plot is displayed below).


```{r}
# find the inspection number from the establishment that ever failed or not
test1 <- inspection %>% mutate (fail=ifelse(results=="Fail",1,0)) %>% group_by(dba_name) %>%summarise(any_fail=if_else(sum(fail)>0,TRUE,FALSE),inspection_number=length(inspection_id)) %>% arrange(desc(inspection_number))
test1 %>% head(5)
```
```{r}
# calculate the percentage of inspections among all
test2 <- test1 %>% mutate(n=100*inspection_number/sum(inspection_number))
test2 %>% head(5)
```

```{r}
# plots
library(mosaicData)
ggplot(data=test2,aes(x=any_fail,y=n)) + geom_boxplot() 
```
# I cannot get the targeted results. I guess I don't understand what kind of distribution need to be analyzed for this question.  


Answers with Python
https://github.com/jamesqiu2021/DSCI-401-QW/blob/main/test1_WQ.ipynb
